---
title: "SPR 2023: Investigating Variation in Effects"
author: "Maria Schweer-Collins, Nicholas Parr, Sean Grant"
date: "2023-05-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# You can use this code to install packages required to follow along
# install.packages(c("robutmeta", "tidyverse", "metafor", "meta", "rio", "here", "readxl"))
library(robumeta)
library(tidyverse)
library(metafor)
library(meta)
library(rio)
library(here)
library(readxl)
```


## Follow along with session "Investigating Variation in Effects"


### Import data
```{r import, echo = TRUE, eval = TRUE}


anxiety_symptoms <- import(here("data", "anxiety_symptoms.csv"))

     
### You should have 16 observations/studies
```

### Run random-effects meta-analysis 

```{r random-effects}
anxiety_re <- rma(yi, vi, data = anxiety_symptoms, method="REML", knha=TRUE)
anxiety_re

# obtain prediction intervals using predict function

predict(anxiety_re)

```


### Visualize the results with a forest plot 

A 95% prediction interval is included in a forest plot by using the addcred = TRUE argument, 
and is denoted by the gray bounding lines around the 
overall effect size estimate (the diamond-shaped indicator). 

```{r forest, out.width= "125%", out.height="165%"}
forest(anxiety_re, 
          header=c("Author and Year", "SMD [95% CI]"), 
          mlab="Random-Effects", 
          order=-yi, 
          slab = anxiety_symptoms$study, 
          xlab = "Standardized Mean Difference",
          addcred = TRUE,
          cex = 1,
          xlim=c(-3.5,3))
text(c(-2,1), -1.75, c("Favors Intervention","Favors Control"), pos=c(4,2), offset=-0.3)
```




### Now we will examine the categorical moderator "comparator_f" using meta-regression

The following model specification fits a mixed-effects meta-regression model to examine 
whether the control group/comparator type moderates intervention effects.
comparator_f is a factor (categorical) variable with three levels: active, treatment as usual (Tx as usual),
and Waitlist of no-treatment control (Control). The treatment as usual comparator serves as the reference level
for this variable. knha = TRUE indicates that the Knapp and Hartung adjustment is used for the omnibus Test of 
Moderators; this adjustment provides a more conservative test of whether included moderators are related to 
the overall effect size estimate.

### First, we investigate our moderator "comparator_f"

```{r check moderator}

## evaluate how many studies exist at each level of the moderator
anxiety_symptoms %>%
  group_by(comparator_f) %>%
  count()

```

There are only two studies with an active comparison group, and all other groups
are relatively small, so we would want to interpret any results from analyses containing
this moderator with caution. 


### Conduct a meta-regression with a three-factor moderator "comparator_f"

```{r meta-regression}
anxiety_comparator <- rma(yi=yi, vi=vi, mods =~ factor(comparator_f),
                          method = "REML", knha = TRUE, data = anxiety_symptoms)
anxiety_comparator

```



### Alternatively, estimate a no intercept model for predicted effects by group

The above mixed-effects model specification provides the effect size estimate and 
corresponding confidence interval for the intercept, which represents the reference 
level of the moderator variable (treatment as usual comparator in this example). 
The additional estimates in the output are interpreted as the difference in effect between 
other comparator types and the reference level (i.e., the values of the coefficients are 
added to or subtracted from the intercept value). In order to examine and report the effect 
estimate and corresponding confidence interval for the other levels of the moderator variable, 
the below model specification should be used. 

```{r meta-regression no intercept}
anxiety_comparator <- rma(yi=yi, vi=vi, mods =~ factor(comparator_f)-1,
                          method = "REML", knha = TRUE, data = anxiety_symptoms)
anxiety_comparator

```


Credit: This tutorial builds from [Parr, Schweer-Collins et al., 2019](https://onlinelibrary.wiley.com/doi/10.1016/j.adolescence.2019.10.009)


      