---
title: "SPR 2023: Meta-Analysis Hands-on Demo in R"
author: "Maria Schweer-Collins"
date: "2023-05-30"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# You can use this code to install packages required to follow along
# install.packages(c("robutmeta", "tidyverse", "metafor", "meta", "rio", "here", "readxl"))
library(robumeta)
library(tidyverse)
library(metafor)
library(meta)
library(rio)
library(here)
library(readxl)
```


## Hands on Demo


### Import depression outcomes data
```{r import, echo = TRUE, eval = TRUE}

depression_symptoms <- import(here("data", "depression_symptoms.csv"))

     
### You should have 40 observations/studies
```

### Run random-effects meta-analysis 

```{r random-effects}
depression_re <- rma(yi, vi, data = depression_symptoms, method="REML", knha=TRUE)
depression_re

# obtain prediction intervals using predict function

predict(depression_re)

```


### Visualize the results with a forest plot 

Forest plots show the study-level and overall effect estimates and the precision with which 
those effects were estimated, but do not typically indicate the range of true effects 
(the prediction interval for the overall effect size estimate). When the prediction interval is 
incorporated, the forest plot presents both sources of variation in the overall effect size 
estimate – within-study error (the confidence interval for each study’s estimated effect) and 
true between-study heterogeneity – and therefore displays a more comprehensive summary of the 
important characteristics of a meta-analysis. A 95% prediction interval is included in a forest 
plot by using the addcred = TRUE argument, and is denoted by the gray bounding lines around the 
overall effect size estimate (the diamond-shaped indicator). The width of the indicator itself 
corresponds to the 95% confidence interval of the overall effect size estimate.

```{r forest, out.width= "125%", out.height="165%"}
forest(depression_re, 
          header=c("Author and Year", "SMD [95% CI]"), 
          mlab="Random-Effects",
          addcred = TRUE,
          order = yi,
          cex = .5,
          xlab = "Standardized Mean Difference",
          slab = depression_symptoms$study)

```



 

### First, we investigate our moderator "measure_f"

measure_f is a four level factor that distinguishes the depression outcome measure, specifically the Beck
Depression Inventory (BDI), the Children's Depression Inventory (CDI), and the Center for Epidemiological StudiesDepression Scale (CES-D), and all other measures (other), inclduing the PHQ-9, RADS, and SMFQ. Some prior research suggests that both the BDI and CES-D are overly sensitive measures of depression, with the
CES-D being more so. The other types of depression measures serve as the reference level
for this variable. knha = TRUE indicates that the Knapp and Hartung adjustment is used for the omnibus Test of 
Moderators; this adjustment provides a more conservative test of whether included moderators are related to 
the overall effect size estimate. In this moderation analysis, we will ask: 

Does the effect of depression prevention programs on depression symptoms vary 
depending on how depression symptoms were measured?

```{r check moderator}

## evaluate how many studies exist at each level of the moderator
depression_symptoms %>%
  group_by(measure_f) %>%
  count()

## graphically we can examine the distribution of the moderator

boxplot(yi ~ measure_f, data = depression_symptoms, xlab = "Depression Measure", 
        ylab = "SMD Effect Size", col=c("pink", "orange", "blue"))

```
 


### Conduct a meta-regression with a three-factor moderator "comparator_f"

```{r meta-regression}
depression_measure <- rma(yi=yi, vi=vi, mods =~ factor(measure_f),
                          method = "REML", knha = TRUE, data = depression_symptoms)
depression_measure

```



### Alternatively, estimate a no intercept model for predicted effects by group

The above mixed-effects model specification provides the effect size estimate and 
corresponding confidence interval for the intercept, which represents the reference 
level of the moderator variable (treatment as usual comparator in this example). 
The additional estimates in the output are interpreted as the difference in effect between 
other comparator types and the reference level (i.e., the values of the coefficients are 
added to or subtracted from the intercept value). In order to examine and report the effect 
estimate and corresponding confidence interval for the other levels of the moderator variable, 
the below model specification should be used. Note the difference between the two specifications 
is the inclusion of - 1 following the moderator variable. 

```{r meta-regression no intercept}
depression_measure <- rma(yi=yi, vi=vi, mods =~ factor(measure_f)-1,
                          method = "REML", knha = TRUE, data = depression_symptoms)
depression_measure

```


### Conduct a meta-regression with primary_prevention

Primary prevention is a two level factor variable that indicates whether
a depression prevention program was delivered as a primary prevention, indicated
through "yes"  (i.e.,before the onset of depression), or secondary 
prevention, indicated by "no" (i.e., prevention delivered to youth with some 
elevations in depression symptoms).


### Are depression prevention programs delivered in schools more effective when they are delivered as primary prevention?


```{r meta-regression prevent}
# specify a no-intercept meta-regression model
depression_prevent <- rma(yi=yi, vi=vi, mods =~ factor(primary_prevention)-1,
                          method = "REML", knha = TRUE, data = depression_symptoms)
depression_prevent

```


### Conduct a meta-regression with a continuous moderator (baseline depression symptoms)

### Research Question: Are the effects of depression prevention programs larger in
samples with higher average baseline depression symptoms?

### Remember this is an aggregate variable

```{r baseline }
depression_baseline <-rma(yi=yi, vi=vi, mods =~ baseline_depression,
                          method = "REML", knha = TRUE, data = depression_symptoms)
depression_baseline
```




### Conduct multivariable meta-regression with categorical and continuous predictor

### Does the effect of depression prevention programs on depression symptoms vary depending on how depression symptoms were measured, when controlling for baseline depression?
```{r multi meta-reg}

depression_multi <- rma(yi=yi, vi=vi, mods =~ baseline_depression + measure_f,
                          method = "REML", knha = TRUE, data = depression_symptoms) 
depression_multi

```

### if time permits, get into groups of two and work together to run one last moderation analysis using meta-regression. Your group can choose to use one of the following: 



### Example 1 - moderator is united_states where "yes" means the study was conducted in the U.S. and a "no means it was conducted outside the U.S.

- Develop a research question
- Assess the overall ombnibus test
- interpret the model


```{r EX1}
mex1 <- rma(yi=yi, vi=vi, mods =~ factor(united_states)-1,
                          method = "REML", knha = TRUE, data = depression_symptoms)
mex1
```

### Example 2 - moderator is secondary_school where "yes" means the study was conducted in secondary schools and "no" means it was conducted in primary schools

```{r EX2}
mex2 <- rma(yi=yi, vi=vi, mods =~ factor(united_states)-1,
                          method = "REML", knha = TRUE, data = depression_symptoms)
mex2
```

### Example 3 - moderator is high_rob risk of bias, where "yes" indicates a study with high risk of bias and "no" means low or unclear risk of bias

```{r EX3}
mex3 <- rma(yi=yi, vi=vi, mods =~ factor(high_rob)-1,
                           method = "REML", knha = TRUE, data = depression_symptoms)
mex3

```




### A bubble plot
```{r bubble plot}
# running a meta-regression with publication year as a continuous moderator 
depression_publication <-rma(yi=yi, vi=vi, mods =~ publication_year,
                          method = "REML", knha = TRUE, data = depression_symptoms)
depression_publication
# Create bubble plot to visualize relationship between predictor and effect size

regplot(depression_publication)

```

Credit: This tutorial builds from [Parr, Schweer-Collins et al., 2019](https://onlinelibrary.wiley.com/doi/10.1016/j.adolescence.2019.10.009)


      